### 2024-03-27 tesis_Ashish/Chapter7/chapter7.tex
update 27 Mar

\chapter{TEPX luminometer for HL-LHC}  %Title of the First Chapter

@@

The Phase II upgrade of the CMS Tracker is a crucial project undertaken to maintain the effectiveness of the CMS experiment during the High Luminosity LHC (HL-LHC) phase, which is expected to start in 2029 \cite{CERN2022}. The HL-LHC will significantly increase the instantaneous luminosity, leading to a higher number of proton-proton collisions per bunch crossing. Instantaneous luminosity will increase to an unprecedented value of $7.5 \times 10^{34} cm^{-2} \: s^{-1}$ which corresponds to 200 proton-proton collisions per bunch crossing. This higher collision rate poses several challenges, including increased radiation levels, higher data rates, and more complex event environment. Run 2 pixel detector will not be able to handle these extreme radiation environment, resolve nearby particle tracks and operate properly to give a reliable estimate of the instantaneous luminosity for such a high pileup. That is why it will be replaced by a new pixel detector. The Phase II upgrade of the CMS tracker comprises of two main components: the Inner Tracker (IT) and the Outer Tracker (OT) \cite{RoyChowdhury:2729279}. The Inner Tracker consists of three main sub-components: the Tracker Endcap Pixel detector (TEPX), the Tracker Forward Pixel detector (TFPX), and the Tracker Barrel Pixel (TBPX) detector as shown in Fig. \ref{fig:Innertracker} \cite{CERN-LHCC-2017-009}.  %The Outer Tracker is responsible for detecting charged particles in the outer region of the CMS detector. It consists of two key subsystems: the silicon strip tracker and the silicon macro pixel tracker \cite{RoyChowdhury:2729279}.

Data Acquisition (DAQ) system for the Phase II CMS Tracker must handle the increased data rates and complexity associated with the HL-LHC. The HL-LHC is expected to achieve a peak luminosity which is around 5-7 times higher than the current LHC design luminosity which the DAQ system must handle. The upgraded CMS Tracker will generate a large amount of data due to the increased granularity of the pixel and strip sensors. The data rates from the tracker modules are expected to be several terabits per second (Tbps) during the HL-LHC operation. The Level-1 Trigger system in the Phase II Upgrade is designed to accept events at a rate of up to 750 kHz. The DAQ system must handle the transmission of trigger primitives to the Level-1 Trigger system and buffer the full event data until the trigger decision is made. %The high-speed optical links used for data transmission between the Front-End Electronics (FEE) and off-detector electronics are expected to operate at data rates of 10-25 Gbps or higher to handle the large data volumes generated by the upgraded tracker.
In the Inner Tracker readout system, control signals are sent to the it at 160 Mb/s to coordinate its detection processes. It, in turn, outputs data at 1.28 Gb/s per chip to the LpGBT, which acts as a data consolidator for up to three chips. This data is then relayed optically to the DTC board, which operates a bidirectional data exchange: 2.5 Gb/s optical down-links channel control information back to the Inner Tracker, while 10 Gb/s up-links carry processed readout data and monitoring details to the DAQ and control system. Moreover, the DTC can interface with a Lumi processor via an optional link, designed for a substantial data transfer rate of approximately 4x25 Gb/s, to handle specialized luminosity data processing. Finally, the DAQ and control system, which receives the comprehensive data at a substantial rate of 25 Gb/s, undertakes the final processing and storage of data.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5 \columnwidth]{ashish_thesis/tepx_DD.png}
  \caption[X-Y view of TEPX double-disk]{ \onehalfspacing View of one TEPX double-disk in the $r-\phi$ plane; individual pixel modules on two planes (double-disk) and five rings are outlined by yellow rectangles. %The planes are separated in z by about 1 cm. The first plane holds modules on both front and back faces, with alternating rings to cover the gaps in r; the second plane holds modules in the same way as the first plane, but with an offset in $\phi$ to cover the gaps in $\phi$.
  }
  \label{fig:Innertracker_40}
\end{figure}

@@

TEPX is designed to function as a high-precision online bunch-by-bunch luminometer, primarily uses pixel cluster counting for luminosity measurement, a method involving the detection and counting of pixel clusters formed by charged particles from the proton collisions. Additionally, TEPX benefits from coincidence counting due to overlapping modules which enhances accuracy and enables systematic studies. %Positioned as part of the forward extension of the Inner Tracker detector,
TEPX features a sophisticated dedicated trigger system for sampling all relevant bunch crossings in an unbiased manner. This system operates at a baseline trigger rate of 75 kHz, which is 10$\%$  of the 750 kHz allocated for physics data. %In terms of data acquisition,
TEPX adapts to different operational conditions: during standard physics runs, it maintains a baseline luminosity trigger rate of 75 kHz, while during van der Meer (vdM) scans, this rate can significantly increase, potentially up to 1 MHz \cite{Collaboration:2706512}. %TEPX readout rate (E-link speed) is estimated to be approximately 1330 Gbps at a pileup of 200  \cite{Collaboration:2706512}.
%The DAQ system of TEPX is designed to handle substantial data rates, with link bandwidth requirements varying from about 79.2 Gb/s at a trigger rate of 825 kHz during nominal physics operations to up to 24 Gb/s at the standard rate of 75 kHz  \cite{Collaboration:2706512}.

TEPX Disk 4 Ring 1 (D4R1) is located in the innermost ring of the last double-disk of the TEPX, is exclusively designed for measuring luminosity and beam-induced background (BIB). D4R1 is optimally placed due to its low radius, large active area, and high z position, making it highly efficient for these measurements. The D4R1 operational model is distinct from the rest of TEPX \cite{Haranko2023}. It employs the entire available trigger rate and readout bandwidth, which varies based on operational mode. This capability ranges from a trigger rate of 825 kHz (similar to the rest of TEPX) up to several MHz, limited by readout bandwidth and the available e-link capacity. In nominal physics conditions, the baseline trigger rate for D4R1 is 825 kHz, while in van der Meer (vdM) conditions, it is 2 MHz, with the potential to reach between 5 and 10 MHz in certain scenarios.

@@

\section{TEPX Luminomter Simulation}

 Simulated data samples for Phase II include full CMS detector description and uses official CMS software (version $\texttt{CMSSW\_10\_6\_0\_patch2}$) which calls GEANT4 for detector simulation \cite{GEANT4:2002zbu}.  Pixel clusters are reconstructed using the standard CMS reconstruction algorithms \cite{cmssw_2024}. It is done by converting raw data to digi format (row, column, charge) and then form clusters from digis. The samples used in simulation use single-neutrino generated events overlaid with a variable number of minimum-bias events (events with any amount of real energy detected in CMS) to simulate different pileup values.  The number of simulated events used in this analysis for each pileup value is shown in Table \ref{tab:12}.

@@

Linearity is one of the systematic uncertainty in the calculation of instantaneous luminosity and PCC visible cross section $\sigma_{vis}$. A linear relation between the mean number of clusters and pileup (PU) imply that $<N_{cl/pp}> = \frac{<N_{cl}>}{\mu} = \frac{\sigma_{vis}}{\sigma_{pp}}$ , $<N_{cl}> =  \frac{\sigma_{vis}}{\sigma_{pp}} \mu $. Linearity indicates that the PCC visible cross section $\sigma_{vis}$ does not depend on the per bunch instantaneous luminosity (ideal luminometer). In ideal scenario, $\sigma_{vis}$ is not dependent on pileup, but this a potential problem with the luminometer. A non-linear relation $<N_{cluster}> = \alpha (PU)^{\gamma}, \gamma \neq 1$ would add non-linear terms in the rate equation $R = \sigma_{vis} L_{inst}$ and cause $\sigma_{vis}$ to vary with the per bunch instantaneous luminosity and pileup values. TEPX luminometer proposed for Phase II shows excellent linearity over entire pileup range.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{ashish_thesis/michigan_2.png}
  \caption[TEPX cluster linear fit per disk]{\onehalfspacing Left: Simulated mean number of clusters for TEPX disks as a function of pileup. A line is fitted between pileup values of 0 and 2, and then extrapolated up to a pileup of 200. Right: %Deviation from linearity for clusters for  TEPX disks. The non-linearity is calculated as
    The relative difference between the data points and the values of the fit function at the respective pileup value. }
  \label{fig:CMS_4205}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{ashish_thesis/michigan_3.png}
  \caption[TEPX clusters linear fit]{\onehalfspacing Left: Simulated mean number of clusters for  TEPX Disk 4 Ring 1 as a function of pileup. A line is fitted between pileup values of 0 and 2, and then extrapolated up to a pileup of 200. Right: %Deviation from linearity for clusters for TEPX Disk 4 Ring 1. The non-linearity is calculated as
    The relative difference between the data points and the values of the fit function at the respective pileup value. }
  \label{fig:CMS_4206}
\end{figure}

@@

A new algorithm for luminosity measurement based on counting two-fold coincidences is proposed that applies selections using dr and $d\phi$ variables (shown in Fig. \ref{fig:drdphi_diagram}) to minimize non-linearity for high pileup as these variables take into account track angle and its curvature \cite{sehrawat2020}.  Proper selections are applied for each TEPX disk based on these variables using low pileup simulated samples and then these selections are applied to high pileup samples for suppressing fake coincidences.

\begin{figure}[!htp]
\centering
\includegraphics[width=0.7\textwidth]{ashish_thesis/dr_dphi_variables1_1.png}
\caption[New Variables For Two Fold Coincidences]{%
  r, phi variables used to define coincidence clusters. %These variables take into account track angle and curvature.
}
\label{fig:drdphi_diagram}
\end{figure}

@@

We study the two-fold coincidences in $\phi$ for all TEPX Disk and rings using r, phi variable for pileup 0.5. Based on pileup 0.5 sample, a pre-selection of 0.15 mm on dr and 0.04 on dphi variable is applied in order to remove fake coincidences.

The dr, dphi distribution showing true, fake, total two-fold coincidences for Disk 4 Ring 1 for pileup 0.5 is shown in Fig. \ref{fig:cluster_dr_dphi_dist}. After obtaining these distributions for all TEPX disks and rings, we carry out fit of the true coincidences distribution (shown in red color) to obtain mean and standard deviation of these distributions. A two-standard-deviation selection is applied on the dr value about the mean. The mean values are shown in Table \ref{tab:mdr_cuts_offset_values}. The  standard deviation values of the fit to true coincidences for all disks and rings are shown in Table \ref{tab:disk_values}. %Example fits are shown in Fig. \ref{fig:cluster_ring_1003}, \ref{fig:cluster_ring_70}, \ref{fig:cluster_ring_71}. 
Similarly, two-sigma cut is applied on the dphi variable as well. The standard deviation values for dphi variable are shown in Table \ref{tab:mdphi_cuts_values}. While doing the fit, mean value for dphi variable is set to zero. %The mean is taken as the offset value and standard deviation is used as the selection on dr, dphi variables as shown in Table \ref{tab:disk_values}, \ref{tab:mdphi_cuts_values} and \ref{tab:mdr_cuts_offset_values}. 
%The mean of the fit is taken as the offset value and standard deviation is used as the cut on dr, dphi variables.
A plot of two-fold coincidences in phi after applying all selection based on standard deviation values of all TEPX disks and rings is shown in Fig. \ref{fig:cluster_ring_720}.

\begin{figure}[H]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/D4R1_S2_drdphi_cut_2.png}
\caption[Pileup 0.5 Cluster Count in D4R1 vs. dr & dphi]{%
  Distribution of cluster count as a function of dr and dphi variables for pileup 0.5 for Disk 4 Ring 1. The black curve represents the total two-fold coincidence counts, the red curve depicts the true two-fold coincidence counts, and the green curve indicates the fake two-fold coincidence counts.
}
\label{fig:cluster_dr_dphi_dist}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/D4R3_S2_drdphi_cut.png}
\caption[Pileup 0.5 Cluster Count in D4R3 (+Z) vs. dr & dphi]{%
  Distribution of cluster count as a function of dr and dphi variables for pileup 0.5 for disk 4 Ring 3 (+Z). Appropriate selection on dr, dphi are used to minimize fake two fold coincidences.
}
\label{fig:cluster_ring_10001}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/D4R5_drdphicut.png}
\caption[Pileup 0.5 Cluster Count in D4R5 (+Z) vs. dr & dphi]{%
  Distribution of cluster count as a function of dr and dphi variables for pileup 0.5 for disk 4 Ring 5 (+Z). Appropriate selection on dr, dphi are used to minimize fake two fold coincidences.
}
\label{fig:cluster_ring_1002}
\end{figure}

@@

\begin{table}[H]
  \centering
  \caption[dr offset values for two-fold in phi]{dr mean values (mm) obtained from fit for two fold coincidences in phi for all disks and rings}
  \begin{tabular}{cccccc}
    \textbf{Disk} & \textbf{Ring 1} & \textbf{Ring 2} & \textbf{Ring 3} & \textbf{Ring 4} & \textbf{Ring 5} \\
    \hline
    %-1 & 0.0342041  & 0.0511416 & 0.0676943 & 0.0834355 & 0.0984932 \\
    %-2 & 0.0298276  & 0.0445398 & 0.0589828 & 0.0727454 & 0.0853587 \\
    %-3 & 0.0260326  & 0.0386608 & 0.05122   & 0.0632478 & 0.0746467 \\
    %-4 & 0.0227089  & 0.0336467 & 0.044476  & 0.0551065 & 0.06484   \\
    1  & 0.034  & 0.051 & 0.068 & 0.084 & 0.098 \\
    2  & 0.030  & 0.044 & 0.059 & 0.073 & 0.085 \\
    3  & 0.026  & 0.039 & 0.051 & 0.063 & 0.074 \\
    4  & 0.023   & 0.034 & 0.045 & 0.055 & 0.064 \\
  \end{tabular}
  \label{tab:mdr_cuts_offset_values}
\end{table}

\begin{table}[H]
  \centering
  \caption[dr cut values for two-fold in phi]{dr cut values (mm) for two fold coincidences in phi for all disks and rings}
  \begin{tabular}{cccccc}
    \textbf{Disk} & \textbf{Ring 1} & \textbf{Ring 2} & \textbf{Ring 3} & \textbf{Ring 4} & \textbf{Ring 5} \\
    \hline
    %-1 & 0.00633554 & 0.00709231 & 0.00757662 & 0.008659   & 0.00872148 \\
    %-2 & 0.00594206 & 0.00663493 & 0.00718173 & 0.00766371 & 0.00797059 \\
    %-3 & 0.00594206 & 0.00663493 & 0.00660603 & 0.00713035 & 0.00679445 \\
    %-4 & 0.00530235 & 0.00587692 & 0.00616465 & 0.00661972 & 0.00660599 \\
    1  & 0.0064 & 0.0072  & 0.0078 & 0.0086 & 0.0088 \\
    2  & 0.0059 & 0.0066 & 0.0070 & 0.0077 & 0.0079 \\
    3  & 0.0056 & 0.0062 & 0.0066 & 0.0070  & 0.0073 \\
    4  & 0.0053  & 0.0059 & 0.0061  & 0.0067 & 0.0068 \\
  \end{tabular}
  \label{tab:disk_values}
\end{table}

\begin{table}[H]
  \centering
  \caption[dphi cut values for two-fold in phi]{dphi cut values for two fold coincidences in phi for all disks and rings}
  \begin{tabular}{cccccc}
    \textbf{Disk} & \textbf{Ring 1} & \textbf{Ring 2} & \textbf{Ring 3} & \textbf{Ring 4} & \textbf{Ring 5} \\
    \hline
    %-1 & 0.000856727  & 0.00128078   & 0.00151297   & 0.00151297   & 0.0016561    \\
    %-2 & 0.000765842  & 0.00110698   & 0.00131104   & 0.00143555   & 0.00150265   \\
    %-3 & 0.000684645  & 0.000975385  & 0.00115899   & 0.00126858   & 0.00131598   \\
    %-4 & 0.000598839  & 0.000862728  & 0.00104327   & 0.00112621   & 0.00115609   \\
    1  & 0.00088  & 0.0013   & 0.0015    & 0.0016   & 0.0017   \\
    2  & 0.00076   & 0.0011   & 0.0013   & 0.0014   & 0.0015   \\
    3  & 0.00068  & 0.00097  & 0.0012   & 0.0013   & 0.0013   \\
    4  & 0.00061  & 0.00088  & 0.0010   & 0.0011   & 0.0012   \\
  \end{tabular}
  \label{tab:mdphi_cuts_values}
\end{table}

@@

\begin{table}[H]
  \centering
  \caption[dphi cut values for two-fold in r]{dphi standard deviation values obtained from fit to true coincidences in r}
\begin{tabular}{ccccc}
 & \multicolumn{4}{c}{\textbf{Ring}} \\
\textbf{Disk} & \textbf{1-2} & \textbf{2-3} & \textbf{3-4} & \textbf{4-5} \\
\hline
%-4 & 0.000628313 & 0.000828659 & 0.000975585 & 0.00101984 \\ 
%-3 & 0.000545699 & 0.000722289 & 0.00081899 & 0.000918094 \\ 
%-2 & 0.000495422 & 0.000649127 & 0.000754224 & 0.000839347 \\ 
%-1 & 0.000451069 & 0.000604104 & 0.000689508 & 0.000755603 \\
1 & 0.00063 & 0.00082 & 0.00094 & 0.0010 \\
2 & 0.00057 & 0.00076 & 0.00085 & 0.00090 \\
3 & 0.00052 & 0.00066 & 0.00077 & 0.00083 \\
4 & 0.00047 & 0.00059 & 0.00069 & 0.00075 \\
\end{tabular}
\label{tab:my_label_54}
\end{table}

>>>>

\chapter{TEPX luminometer for HL-LHC}  %Title of the First Chapter
\label{TEPX luminometer for HL-LHC}

@@

The Phase II upgrade of the CMS Tracker is a crucial project undertaken to maintain the effectiveness of the CMS experiment during the High Luminosity LHC (HL-LHC) phase, which is expected to start in 2029 \cite{CERN2022}. The HL-LHC will significantly increase the instantaneous luminosity, leading to a higher number of proton-proton collisions per bunch crossing. Instantaneous luminosity will increase to an unprecedented value of $7.5 \times 10^{34} cm^{-2} \: s^{-1}$ which corresponds to 200 proton-proton collisions per bunch crossing, on average. This higher collision rate poses several challenges, including increased radiation levels, higher data rates, and more complex event environment. The Run 2 pixel detector will not be able to handle these extreme radiation environment, resolve nearby particle tracks and operate properly to give a reliable estimate of the instantaneous luminosity for such a high pileup. That is why it will be replaced by a new pixel detector. The Phase II upgrade of the CMS tracker comprises of two main components: the Inner Tracker (IT) and the Outer Tracker (OT) \cite{RoyChowdhury:2729279}. The Inner Tracker consists of three main sub-components: the Tracker Endcap Pixel detector (TEPX), the Tracker Forward Pixel detector (TFPX), and the Tracker Barrel Pixel (TBPX) detector as shown in Fig. \ref{fig:Innertracker} \cite{CERN-LHCC-2017-009}.  %The Outer Tracker is responsible for detecting charged particles in the outer region of the CMS detector. It consists of two key subsystems: the silicon strip tracker and the silicon macro pixel tracker \cite{RoyChowdhury:2729279}.

The Data Acquisition (DAQ) system for the Phase II CMS Tracker must handle the increased data rates and complexity associated with the HL-LHC. The HL-LHC is expected to achieve a peak luminosity which is around 5-7 times higher than the current LHC design luminosity which the DAQ system must handle. The upgraded CMS Tracker will generate a large amount of data due to the increased granularity of the pixel and strip sensors. The data rates from the tracker modules are expected to be several terabits per second (Tbps) during the HL-LHC operation. The Level-1 Trigger system in the Phase II Upgrade is designed to accept events at a rate of up to 750 kHz. The DAQ system must handle the transmission of trigger primitives to the Level-1 Trigger system and buffer the full event data until the trigger decision is made. %The high-speed optical links used for data transmission between the Front-End Electronics (FEE) and off-detector electronics are expected to operate at data rates of 10-25 Gbps or higher to handle the large data volumes generated by the upgraded tracker.
In the Inner Tracker readout system, control signals are sent to the inner tracker at 160 Mb/s to coordinate its detection processes. It, in turn, outputs data at 1.28 Gb/s per chip to the LpGBT, which acts as a data consolidator for up to three chips. This data is then relayed optically to the DTC board, which operates a bidirectional data exchange: 2.5 Gb/s optical down-links channel control information back to the Inner Tracker, while 10 Gb/s up-links carry processed readout data and monitoring details to the DAQ and control system. Moreover, the DTC can interface with a Lumi processor via an optional link, designed for a substantial data transfer rate of approximately 4x25 Gb/s, to handle specialized luminosity data processing. Finally, the DAQ and control system, which receives the comprehensive data at a substantial rate of 25 Gb/s, undertakes the final processing and storage of data.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5 \columnwidth]{ashish_thesis/tepx_DD.png}
  \caption[X-Y view of TEPX double-disk]{ \onehalfspacing View of one TEPX double-disk in the $r-\phi$ plane; individual pixel modules on two planes (double-disk) and five rings are outlined by orange rectangles \cite{Collaboration:275907420}. %The planes are separated in z by about 1 cm. The first plane holds modules on both front and back faces, with alternating rings to cover the gaps in r; the second plane holds modules in the same way as the first plane, but with an offset in $\phi$ to cover the gaps in $\phi$.
  }
  \label{fig:Innertracker_40}
\end{figure}

@@

TEPX is designed to function as a high-precision online bunch-by-bunch luminometer, primarily using pixel cluster counting for luminosity measurement, a method involving the detection and counting of pixel clusters formed by charged particles from the proton collisions. Additionally, TEPX benefits from coincidence counting due to overlapping modules which enhances accuracy and enables systematic studies. %Positioned as part of the forward extension of the Inner Tracker detector,
TEPX features a sophisticated dedicated trigger system for sampling all relevant bunch crossings in an unbiased manner. This system operates at a baseline trigger rate of 75 kHz, which is 10$\%$  of the 750 kHz allocated for physics data. %In terms of data acquisition,
TEPX adapts to different operational conditions: during standard physics runs, it maintains a baseline luminosity trigger rate of 75 kHz, while during vdM scans, this rate can significantly increase, potentially up to 1 MHz \cite{Collaboration:2706512}. %TEPX readout rate (E-link speed) is estimated to be approximately 1330 Gbps at a pileup of 200  \cite{Collaboration:2706512}.
%The DAQ system of TEPX is designed to handle substantial data rates, with link bandwidth requirements varying from about 79.2 Gb/s at a trigger rate of 825 kHz during nominal physics operations to up to 24 Gb/s at the standard rate of 75 kHz  \cite{Collaboration:2706512}.

TEPX Disk 4 Ring 1 (D4R1) is located in the innermost ring of the last double-disk of the TEPX and it is exclusively designed for measuring luminosity and beam-induced background (BIB). D4R1 is optimally placed due to its low radius, large active area, and high z position, making it highly efficient for these measurements. The D4R1 operational mode is distinct from the rest of TEPX \cite{Haranko2023}. It employs the entire available trigger rate and readout bandwidth, which varies based on operational mode. This capability ranges from a trigger rate of 825 kHz (similar to the rest of TEPX) up to several MHz, limited by readout bandwidth and the available e-link capacity. In nominal physics conditions, the baseline trigger rate for D4R1 is 825 kHz, while in van der Meer (vdM) conditions, it is 2 MHz, with the potential to reach between 5 and 10 MHz in certain scenarios.

@@

\section{TEPX Luminometer Simulation}

 Simulated data samples for Phase II include full CMS detector description and use official CMS software (version $\texttt{CMSSW\_10\_6\_0\_patch2}$) which calls GEANT4 for detector simulation \cite{GEANT4:2002zbu}.  Pixel clusters are reconstructed using the standard CMS reconstruction algorithms \cite{cmssw_2024}. It is done by converting raw data to digi format (row, column, charge) and then form clusters from digis. The samples used in simulation use single-neutrino generated events overlaid with a variable number of minimum-bias events (events with any amount of real energy detected in CMS) to simulate different pileup values.  The number of simulated events used in this analysis for each pileup value is shown in Table \ref{tab:12}.

@@

Linearity is one of the systematic uncertainties in the calculation of instantaneous luminosity and PCC visible cross section $\sigma_{vis}$. A linear relation between the mean number of clusters and pileup (PU) imply that $<N_{cl/pp}> = \frac{<N_{cl}>}{\mu} = \frac{\sigma_{vis}}{\sigma_{pp}}$ , $<N_{cl}> =  \frac{\sigma_{vis}}{\sigma_{pp}} \mu $. Linearity indicates that the PCC visible cross section $\sigma_{vis}$ does not depend on the per bunch instantaneous luminosity (ideal luminometer). In ideal scenario, $\sigma_{vis}$ is not dependent on pileup, but this a potential problem with the luminometer. A non-linear relation $<N_{cluster}> = \alpha (PU)^{\gamma}, \gamma \neq 1$ would add non-linear terms in the rate equation $R = \sigma_{vis} L_{inst}$ and cause $\sigma_{vis}$ to vary with the per bunch instantaneous luminosity and pileup values. TEPX luminometer proposed for Phase II shows excellent linearity over entire pileup range.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{ashish_thesis/michigan_2.png}
  \caption[TEPX cluster linear fit per disk]{\onehalfspacing Left: Simulated mean number of clusters for TEPX disks as a function of pileup. A line is fitted between pileup values of 0 and 2, and then extrapolated up to a pileup of 200. Right: %Deviation from linearity for clusters for  TEPX disks. The non-linearity is calculated as
    The relative difference between the data points and the values of the fit function at the respective pileup value \cite{Collaboration:275907420}.}
  \label{fig:CMS_4205}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=1\columnwidth]{ashish_thesis/michigan_3.png}
  \caption[TEPX clusters linear fit]{\onehalfspacing Left: Simulated mean number of clusters for  TEPX Disk 4 Ring 1 as a function of pileup. A line is fitted between pileup values of 0 and 2, and then extrapolated up to a pileup of 200. Right: %Deviation from linearity for clusters for TEPX Disk 4 Ring 1. The non-linearity is calculated as
    The relative difference between the data points and the values of the fit function at the respective pileup value \cite{Collaboration:275907420}. }
  \label{fig:CMS_4206}
\end{figure}

@@

A new algorithm for luminosity measurement based on counting two-fold coincidences is proposed that applies selections using dr and d$\phi$ variables (shown in Fig. \ref{fig:drdphi_diagram}) to minimize non-linearity for high pileup as these variables take into account track angle and its curvature \cite{sehrawat2020}.  Proper selections are applied for each TEPX disk based on these variables using low pileup simulated samples and then these selections are applied to high pileup samples for suppressing fake coincidences.

\begin{figure}[!htp]
\centering
\includegraphics[width=0.7\textwidth]{ashish_thesis/dr_dphi_variables1_1.png}
\caption[New Variables For Two Fold Coincidences]{%
  r, $\phi$ variables used to define coincidence clusters. %These variables take into account track angle and curvature.
}
\label{fig:drdphi_diagram}
\end{figure}

@@

We study the two-fold coincidences in $\phi$ for all TEPX Disk and rings using r, $\phi$ variable for pileup 0.5. Based on pileup 0.5 sample, a pre-selection of 0.15 mm on dr and 0.04 on d$\phi$ variable is applied in order to remove fake coincidences.

The dr, d$\phi$ distribution showing true, fake, total two-fold coincidences for Disk 4 Ring 1 for pileup 0.5 is shown in Fig. \ref{fig:cluster_dr_dphi_dist}. After obtaining these distributions for all TEPX disks and rings, we carry out a fit of the true coincidences distribution (shown in red color) to obtain mean and standard deviation of these distributions. A two-standard-deviation selection is applied on the dr value about the mean. The mean values are shown in Table \ref{tab:mdr_cuts_offset_values}. The  standard deviation values of the fit to true coincidences for all disks and rings are shown in Table \ref{tab:disk_values}. %Example fits are shown in Fig. \ref{fig:cluster_ring_1003}, \ref{fig:cluster_ring_70}, \ref{fig:cluster_ring_71}. 
Similarly, two-sigma cut is applied on the d$\phi$ variable as well. The standard deviation values for d$\phi$ variable are shown in Table \ref{tab:mdphi_cuts_values}. While doing the fit, the mean value for d$\phi$ variable is set to zero. %The mean is taken as the offset value and standard deviation is used as the selection on dr, dphi variables as shown in Table \ref{tab:disk_values}, \ref{tab:mdphi_cuts_values} and \ref{tab:mdr_cuts_offset_values}. 
%The mean of the fit is taken as the offset value and standard deviation is used as the cut on dr, dphi variables.
A plot of two-fold coincidences in $\phi$ after applying all selection based on standard deviation values of all TEPX disks and rings is shown in Fig. \ref{fig:cluster_ring_720}.

\begin{figure}[H]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/D4R1_S2_drdphi_cut_2.png}
\caption[Pileup 0.5 Cluster Count in D4R1 vs. dr & d$\phi$]{%
  Distribution of cluster count as a function of dr and d$\phi$ variables for pileup 0.5 for Disk 4 Ring 1. The black curve represents the total two-fold coincidence counts, the red curve depicts the true two-fold coincidence counts, and the green curve indicates the fake two-fold coincidence counts.
}
\label{fig:cluster_dr_dphi_dist}
\end{figure}

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/D4R3_S2_drdphi_cut.png}
\caption[Pileup 0.5 Cluster Count in D4R3 (+Z) vs. dr & d$\phi$]{%
  Distribution of cluster count as a function of dr and dphi variables for pileup 0.5 for disk 4 Ring 3 (+Z). Appropriate selection on dr, dphi are used to minimize fake two fold coincidences.
}
\label{fig:cluster_ring_10001}
\end{figure}

\begin{figure}[!htp]
\centering
\includegraphics[width=1\textwidth]{ashish_thesis/D4R5_drdphicut.png}
\caption[Pileup 0.5 Cluster Count in D4R5 (+Z) vs. dr & d$\phi$]{%
  Distribution of cluster count as a function of dr and dphi variables for pileup 0.5 for disk 4 Ring 5 (+Z). Appropriate selection on dr, dphi are used to minimize fake two fold coincidences.
}
\label{fig:cluster_ring_1002}
\end{figure}

@@

\begin{table}[H]
  \centering
  \caption[dr offset values for two-fold in $\phi$]{dr mean values (mm) obtained from fit for two fold coincidences in $\phi$ for all disks and rings}
  \begin{tabular}{cccccc}
    \textbf{Disk} & \textbf{Ring 1} & \textbf{Ring 2} & \textbf{Ring 3} & \textbf{Ring 4} & \textbf{Ring 5} \\
    \hline
    %-1 & 0.0342041  & 0.0511416 & 0.0676943 & 0.0834355 & 0.0984932 \\
    %-2 & 0.0298276  & 0.0445398 & 0.0589828 & 0.0727454 & 0.0853587 \\
    %-3 & 0.0260326  & 0.0386608 & 0.05122   & 0.0632478 & 0.0746467 \\
    %-4 & 0.0227089  & 0.0336467 & 0.044476  & 0.0551065 & 0.06484   \\
    1  & 0.034  & 0.051 & 0.068 & 0.084 & 0.098 \\
    2  & 0.030  & 0.044 & 0.059 & 0.073 & 0.085 \\
    3  & 0.026  & 0.039 & 0.051 & 0.063 & 0.074 \\
    4  & 0.023   & 0.034 & 0.045 & 0.055 & 0.064 \\
  \end{tabular}
  \label{tab:mdr_cuts_offset_values}
\end{table}

\begin{table}[H]
  \centering
  \caption[dr cut values for two-fold in $\phi$]{dr cut values (mm) for two fold coincidences in $\phi$ for all disks and rings}
  \begin{tabular}{cccccc}
    \textbf{Disk} & \textbf{Ring 1} & \textbf{Ring 2} & \textbf{Ring 3} & \textbf{Ring 4} & \textbf{Ring 5} \\
    \hline
    %-1 & 0.00633554 & 0.00709231 & 0.00757662 & 0.008659   & 0.00872148 \\
    %-2 & 0.00594206 & 0.00663493 & 0.00718173 & 0.00766371 & 0.00797059 \\
    %-3 & 0.00594206 & 0.00663493 & 0.00660603 & 0.00713035 & 0.00679445 \\
    %-4 & 0.00530235 & 0.00587692 & 0.00616465 & 0.00661972 & 0.00660599 \\
    1  & 0.0064 & 0.0072  & 0.0078 & 0.0086 & 0.0088 \\
    2  & 0.0059 & 0.0066 & 0.0070 & 0.0077 & 0.0079 \\
    3  & 0.0056 & 0.0062 & 0.0066 & 0.0070  & 0.0073 \\
    4  & 0.0053  & 0.0059 & 0.0061  & 0.0067 & 0.0068 \\
  \end{tabular}
  \label{tab:disk_values}
\end{table}

\begin{table}[H]
  \centering
  \caption[d$\phi$ cut values for two-fold in $\phi$]{d$\phi$ cut values for two fold coincidences in $\phi$ for all disks and rings}
  \begin{tabular}{cccccc}
    \textbf{Disk} & \textbf{Ring 1} & \textbf{Ring 2} & \textbf{Ring 3} & \textbf{Ring 4} & \textbf{Ring 5} \\
    \hline
    %-1 & 0.000856727  & 0.00128078   & 0.00151297   & 0.00151297   & 0.0016561    \\
    %-2 & 0.000765842  & 0.00110698   & 0.00131104   & 0.00143555   & 0.00150265   \\
    %-3 & 0.000684645  & 0.000975385  & 0.00115899   & 0.00126858   & 0.00131598   \\
    %-4 & 0.000598839  & 0.000862728  & 0.00104327   & 0.00112621   & 0.00115609   \\
    1  & 0.00088  & 0.0013   & 0.0015    & 0.0016   & 0.0017   \\
    2  & 0.00076   & 0.0011   & 0.0013   & 0.0014   & 0.0015   \\
    3  & 0.00068  & 0.00097  & 0.0012   & 0.0013   & 0.0013   \\
    4  & 0.00061  & 0.00088  & 0.0010   & 0.0011   & 0.0012   \\
  \end{tabular}
  \label{tab:mdphi_cuts_values}
\end{table}

@@

\begin{table}[H]
  \centering
  \caption[d$\phi$ cut values for two-fold in r]{d$\phi$ standard deviation values obtained from fit to true coincidences in r}
\begin{tabular}{ccccc}
 & \multicolumn{4}{c}{\textbf{Ring}} \\
\textbf{Disk} & \textbf{1-2} & \textbf{2-3} & \textbf{3-4} & \textbf{4-5} \\
\hline
%-4 & 0.000628313 & 0.000828659 & 0.000975585 & 0.00101984 \\ 
%-3 & 0.000545699 & 0.000722289 & 0.00081899 & 0.000918094 \\ 
%-2 & 0.000495422 & 0.000649127 & 0.000754224 & 0.000839347 \\ 
%-1 & 0.000451069 & 0.000604104 & 0.000689508 & 0.000755603 \\
1 & 0.00063 & 0.00082 & 0.00094 & 0.0010 \\
2 & 0.00057 & 0.00076 & 0.00085 & 0.00090 \\
3 & 0.00052 & 0.00066 & 0.00077 & 0.00083 \\
4 & 0.00047 & 0.00059 & 0.00069 & 0.00075 \\
\end{tabular}
\label{tab:my_label_54}
\end{table}
